%! suppress = EscapeUnderscore
\chapter{Constrained Optimization} \label{ch:constrained-optimization}

Constrained optimization is the problem of finding a minimum (or maximum) of a function $f(x)$ called the
``objective function'', subject to a number of constraints of the following types:
\bit
\item $h_{i}(x) =0, \:\:\: i=1,2,\ldots$ called ``equality constraints''
\item $g_{i}(x) \leq 0, \:\:\: i=1,2,\ldots$ called ``inequality constraints''
\eit

Let's start first with the equality constraints, and then we will add inequality constraints.

\section{Equality Constrained Optimization}

The formulation of the optimization problem is to optimize $f(x)$ subject to $h_{i}(x) = 0, \:\:\: i=1,2,\ldots$.
Let's assume for simplicity only one constraint $h_{1}(x) = h(x) = 0$. The idea here is that the point that $f(x)$
touches $h(x)$ is the point that $f(x)$ is minimum (or maximum) while the constraint is also valid. At that point $f
(x)$ is parallel to $h(x)$ and the tangents $\nabla_{\boldsymbol{w}} f(x)$ and $\nabla_{\boldsymbol{w}} h(x)$ which
are perpendicular to $f(x)$ and $h(x)$ respectively, are also parallel to each other. Hence, since
$\nabla_{\boldsymbol{w}} f(x)$ and $\nabla_{\boldsymbol{w}} h(x)$ are parallel this translates to:
\bse
\nabla_{\boldsymbol{w}} f(x) = \mu \nabla_{\boldsymbol{w}} h(x)
\ese

\vspace{3pt}

which is the condition for the $f(x)$ to be minimum (or maximum) while $h(x) = 0$. \v

Without loss of generality the condition for many constraints reads:
\bse
\nabla_{\boldsymbol{w}} f(x) = \sum_{i} \mu_{i} \nabla_{\boldsymbol{w}} h_{i}(x)
\ese

or by bringing everything in one side:
\begin{align*}
& \nabla_{\boldsymbol{w}} f(x) - \sum_{i} \mu_{i} \nabla_{\boldsymbol{w}} h_{i}(x) = 0 \\
& \nabla_{\boldsymbol{w}} (f(x) - \sum_{i} \mu_{i} h_{i}(x)) = 0
\end{align*}

At this point we define the ``Lagrangian'' as follows:
\bse
\mathcal{L}(x, \mu_{i}) = f(x) + \sum_{i} \mu_{i} h_{i}(x)
\ese

where $\mu_{i}$ are called ``Lagrange multipliers''. Subsequently, the necessary conditions for optimization of $
\mathcal{L}$ turns to:
\bse
\nabla_{\boldsymbol{w}} \mathcal{L} = 0 \:\:\: \text{and} \:\:\: \frac{\partial \mathcal{L}}{\partial \mu_{i}} = 0
\ese

The solution of this system of equations minimizes (or maximizes) $f(x)$ subject to $h_{i}(x) = 0, \:\:\: \forall i$.

\section{Equality \& Inequality Constrained Optimization}

Now on top of equality constraints we also have inequality constraints. The formulation of the optimization problem
is to optimize $f(x)$ subject to $h_{i}(x) = 0, \:\:\: i=1,2,\ldots$ and $g_{i}(x) \leq 0, \:\:\: i=1,2,\ldots$.
Following a similar way of thinking as before, although a bit more technical, we can show (but we won't) that if the
following four conditions, called ``Karush - Kuhn - Talker conditions'' (KKT), hold:
\bit
\item $h_{i}(x) = 0, \:\:\: \forall i$
\item $g_{i}(x) \leq 0, \:\:\: \forall i$
\item $\lambda_{i} \leq 0, \:\:\: \forall i$
\item $\lambda_{i} g_{i}(x) = 0, \:\:\: \forall i$
\eit

then there exist constants $\mu_{i}$ and $\lambda_{i}$ called ``KKT multipliers'' such that:
\bse
\nabla_{\boldsymbol{w}} f(x) =
\sum_{i} \mu_{i} \nabla_{\boldsymbol{w}} h_{i} (x) + \sum_{i} \lambda_{i} \nabla_{\boldsymbol{w}} g_{i}(x)
\ese

By following the same philosophy as for the equality constrained optimization we define the ``Lagrangian'' as follows:
\bse
\mathcal{L}(x, \mu_{i}, \lambda_{i}) = f(x) + \sum_{i} \mu_{i} h_{i} (x) + \sum_{i} \lambda_{i} g_{i}(x)
\ese

and the necessary conditions for optimization of $ \mathcal{L}$ turn to:
\bse
\nabla_{\boldsymbol{w}} \mathcal{L}
= 0 \:\:\: \text{and} \:\:\: \frac{\partial \mathcal{L}}{\partial \mu_{i}}
= 0 \:\:\: \text{and} \:\:\: \frac{\partial \mathcal{L}}{\partial \lambda_{i}}
= 0
\ese

This final form of the optimization problem is usually called the ``dual optimization problem''. \v

This is the most general case of constrained optimization. If there are no equality constraints $h_{i}(x)$ then we
simply have a theory for inequality constrained optimization. If there are no inequality constraints $g_{i}(x)$ then
the whole theory turns to the equality constrained optimization problem we developed previously and the KKT
multipliers turn to Lagrangian multipliers. Finally, if there are no equality constraints $h_{i}(x)$ neither
inequality constraints $g_{i}(x)$ the theory is just a usual optimization problem where we just find the solution
where the derivative of $f(x)$ is zero.

\chapter{Kernels} \label{ch:kernels}

\bd [Kernel]
Let $\bar{x}$ and $\bar{x}^\prime$ be two vectors of space $X$ and $\Phi$ a non-linear transformation. We define
$\bar{z}$ and $\bar{z}^\prime$ as the transformed vectors $\Phi(\bar{x})$ and $\Phi(\bar{x}^\prime)$:
\begin{align*}
& \bar{x} \in X \xrightarrow{\text{$\Phi$}} \bar{z} = \Phi(\bar{x}) \in Z \\
& \bar{x}^\prime \in X \xrightarrow{\text{$\Phi$}} \bar{z}^\prime = \Phi (\bar{x}^\prime) \in Z
\end{align*}

We define the \textbf{kernel} of space $Z$ as the function that is equal to the inner product of the transformation
vectors:
\bse
K(\bar{x}, \bar{x}^\prime) = \bar{z}^T \bar{z}^\prime
\ese
\ed

Let us for example assume the following non-linear transformation:
\begin{align*}
& \bar{x} = (x_1, x_2) \xrightarrow{\text{$\Phi$}} \bar{z} = \Phi (\bar{x}) = (1, x_1^2, x_2^2, \sqrt{2} x_1,
\sqrt{2} x_2, \sqrt{2} x_1 x_2) \\
& \bar{x}^\prime = (x_1^\prime, x_2^\prime) \xrightarrow{\text{$\Phi$}} \bar{z}^\prime = \Phi(\bar{x}^\prime) =
(1, x_1^{2^\prime}, x_2^{2^\prime}, \sqrt{2} x_1^\prime, \sqrt{2} x_2^\prime, \sqrt{2} x_1^\prime x_2^\prime)
\end{align*}

Then for the inner product:
\bse
\bar{z}^T \bar{z}^\prime = 1 + x_1^2 x_1^{2^\prime} + x_2^2 x_2^{2^\prime} + 2 x_1 x_1^\prime + 2 x_2 x_2^\prime + 2
x_1 x_2 x_1^\prime x_2^\prime
\ese

However, we can get to the same result by simply defining a Kernel of the form:
\begin{align*}
K(\bar{x}, \bar{x}^\prime) &= (1 + \bar{x} \bar{x}^\prime)^2 \\
&= (1 + x_1 x_1^\prime + x_2 x_2^\prime )^2 \\
&= 1 + x_1^2 x_1^{2^\prime} + x_2^2 x_2^{2^\prime} + 2 x_1 x_1^\prime + 2 x_2 x_2^\prime
+ 2 x_1 x_2 x_1^\prime x_2^\prime
\end{align*}

Hence, by knowing the Kernel of a space $Z$ of some non-linear transformation $\Phi$ we can compute inner products
without the need of transforming vectors from $X$ to $Z$. \v

The kernel trick is to use this idea in the opposite direction. Namely, to assume that a function $K(\bar{x},
\bar{x}^\prime)$ is the kernel of some space $Z$ for some non-linear transformation $\Phi$ and to compute inner
products without even knowing the transformation. \v

The question that arises is how do we know that some function $K(\bar{x}, \bar{x}^\prime)$ is actually the kernel of
a space $Z$. There are three approaches to this problem:
\ben
\item By construction (as we did in the example above).
\item By Mercer's condition that states that $K(\bar{x}, \bar{x}^\prime)$ is a valid kernel for some space $Z$ if
\bse
\int K(\bar{x}, \bar{x}^\prime) g(\bar{x}) g(\bar{x}^\prime) d\bar{x} d\bar{x}^\prime \geq 0 \:\:\: \forall \:\:\:
\text{square integrable functions} \:\:\: g(\bar{x})
\ese

\item Sometimes we don't care if $K(\bar{x}, \bar{x}^\prime)$ is a valid kernel for some space $Z$ as long as it does
the job.
\een

\chapter{Clean Code}

Code is clean if it can be understood easily by everyone on the team. Clean code can be read and enhanced by a developer
other than its original author. With understandability comes readability, changeability, extensibility and
maintainability. \v

General Rules:
\bit
\item Follow standard conventions.
\item Keep it simple stupid. Simpler is always better. Reduce complexity as much as possible.
\item Boy scout rule. Leave the campground cleaner than you found it.
\item Always find root cause. Always look for the root cause of a problem.
\eit

Design Rules:
\bit
\item Keep configurable data at high levels.
\item Prefer polymorphism to if/else or switch/case.
\item Separate multi-threading code.
\item Prevent over-configurability.
\item Use dependency injection.
\item Follow Law of Demeter. A class should know only its direct dependencies.
\eit

Understandability Tips:
\bit
\item Be consistent. If you do something a certain way, do all similar things in the same way.
\item Use explanatory variables.
\item Encapsulate boundary conditions. Boundary conditions are hard to keep track of. Put the processing for them in
one place.
\item Prefer dedicated value objects to primitive type.
\item Avoid logical dependency. Don't write methods which works correctly depending on something else in the same class.
\item Avoid negative conditionals.
\eit

Names Rules:
\bit
\item Choose descriptive and unambiguous names.
\item Make meaningful distinction.
\item Use pronounceable names.
\item Use searchable names.
\item Replace magic numbers with named constants.
\item Avoid encodings. Don't append prefixes or type information.
\eit

Functions Rules:
\bit
\item Small.
\item Do one thing.
\item Use descriptive names.
\item Prefer fewer arguments.
\item Have no side effects.
\item Don't use flag arguments. Split method into several independent methods that can be called from the client without
the flag.
\eit

Comments Rules:
\bit
\item Always try to explain yourself in code.
\item Don't be redundant.
\item Don't add obvious noise.
\item Don't use closing brace comments.
\item Don't comment out code. Just remove.
\item Use as explanation of intent.
\item Use as clarification of code.
\item Use as warning of consequences.
\eit

Source Code Structure:
\bit
\item Separate concepts vertically.
\item Related code should appear vertically dense.
\item Declare variables close to their usage.
\item Dependent functions should be close.
\item Similar functions should be close.
\item Place functions in the downward direction.
\item Keep lines short.
\item Don't use horizontal alignment.
\item Use white space to associate related things and disassociate weakly related.
\item Don't break indentation.
\eit

Objects \& Data Structures:
\bit
\item Hide internal structure.
\item Prefer data structures.
\item Avoid hybrids structures (half object and half data).
\item Should be small.
\item Do one thing.
\item Small number of instance variables.
\item Base class should know nothing about their derivatives.
\item Better to have many functions than to pass some code into a function to select a behavior.
\item Prefer non-static methods to static methods.
\eit

Tests:
\bit
\item One assert per test.
\item Readable.
\item Fast.
\item Independent.
\item Repeatable.
\eit

Code Smells:
\bit
\item Rigidity. The software is difficult to change. A small change causes a cascade of subsequent changes.
\item Fragility. The software breaks in many places due to a single change.
\item Immobility. You cannot reuse parts of the code in other projects because of involved risks and high effort.
\item Needless Complexity.
\item Needless Repetition.
\item Opacity. The code is hard to understand.
\eit

\chapter{Software Engineering At Google}

What Is Software Engineering:
\bit
\item Software engineering differs from programming in dimensionality: programming is about producing code. Software
engineering extends that to include the maintenance of that code for its useful life span.
\item There is a factor of at least 100,000 times between the life spans of short-lived code and long-lived code. It is
silly to assume that the same best practices apply universally on both ends of that spectrum.
\item Software is sustainable when, for the expected life span of the code, we are capable of responding to changes in
dependencies, technology, or product requirements. We may choose to not change things, but we need to be capable.
\item Hyrum's Law: with a sufficient number of users of an API, it does not matter what you promise in the contract: all
observable behaviors of your system will be depended on by somebody.
\item Every task your organization has to do repeatedly should be scalable (linear or better) in terms of human input.
Policies are a wonderful tool for making process scalable.
\item Process inefficiencies and other software-development tasks tend to scale up slowly. Be careful about boiled-frog
problems.
\item Expertise pays off particularly well when combined with economies of scale.
\item "Because I said so" is a terrible reason to do things.
\item Being data driven is a good start, but in reality, most decisions are based on a mix of data, assumption,
precedent, and argument. It's best when objective data makes up the majority of those inputs, but it can rarely be all
of them.
\item Being data driven over time implies the need to change directions when the data changes (or when assumptions are
dispelled). Mistakes or revised plans are inevitable.
\eit

How To Work Well On Teams:
\bit
\item Be aware of the trade-offs of working in isolation.
\item Acknowledge the amount of time that you and your team spend communicating and in interpersonal conflict. A small
investment in understanding personalities and working styles of yourself and others can go a long way toward improving
productivity.
\item If you want to work effectively with a team or a large organization, be aware of your preferred working style and
that of others.
\eit

Knowledge Sharing:
\bit
\item Psychological safety is the foundation for fostering a knowledge-sharing environment.
\item Start small: ask questions and write things down.
\item Make it easy for people to get the help they need from both human experts and documented references.
\item At a systemic level, encourage and reward those who take time to teach and broaden their expertise beyond just
themselves, their team, or their organization.
\item There is no silver bullet: empowering a knowledge-sharing culture requires a combination of multiple strategies,
and the exact mix that works best for your organization will likely change over time.
\eit

Engineering For Equity:
\bit
\item Bias is the default.
\item Diversity is necessary to design properly for a comprehensive user base.
\item Inclusivity is critical not just to improving the hiring pipeline for underrepresented groups, but to providing a
truly supportive work environment for all people.
\item Product velocity must be evaluated against providing a product that is truly useful to all users. It's better to
slow down than to release a product that might cause harm to some users.
\eit

How To Lead A Team:
\bit
\item Don't "manage" in the traditional sense; focus on leadership, influence, and serving your team.
\item Delegate where possible; don't DIY (Do It Yourself).
\item Pay particular attention to the focus, direction, and velocity of your team.
\eit

Leading At Scale:
\bit
\item Always Be Deciding: Ambiguous problems have no magic answer; they're all about finding the right trade-offs of the
moment, and iterating.
\item Always Be Leaving: Your job, as a leader, is to build an organization that automatically solves a class of
ambiguous problems—over time—without you needing to be present.
\item Always Be Scaling: Success generates more responsibility over time, and you must proactively manage the scaling
of this work in order to protect your scarce resources of personal time, attention, and energy.
\eit

Measuring Engineering Productivity:
\bit
\item Before measuring productivity, ask whether the result is actionable, regardless of whether the result is positive
or negative. If you can't do anything with the result, it is likely not worth measuring.
\item Select meaningful metrics using the GSM framework. A good metric is a reasonable proxy to the signal you're trying
to measure, and it is traceable back to your original goals.
\item Select metrics that cover all parts of productivity (QUANTS). By doing this, you ensure that you aren't improving
one aspect of productivity (like developer velocity) at the cost of another (like code quality).
\item Qualitative metrics are metrics, too! Consider having a survey mechanism for tracking longitudinal metrics about
engineers' beliefs. Qualitative metrics should also align with the quantitative metrics; if they do not, it is likely
the quantitative metrics that are incorrect.
\item Aim to create recommendations that are built into the developer workflow and incentive structures. Even though it
is sometimes necessary to recommend additional training or documentation, change is more likely to occur if it is built
into the developer's daily habits.
\eit

Style Guides And Rules:
\bit
\item Rules and guidance should aim to support resilience to time and scaling.
\item Know the data so that rules can be adjusted.
\item Not everything should be a rule.
\item Consistency is key.
\item Automate enforcement when possible.
\eit

Code Review:
\bit
\item Code review has many benefits, including ensuring code correctness, comprehension, and consistency across a
codebase.
\item Always check your assumptions through someone else; optimize for the reader.
\item Provide the opportunity for critical feedback while remaining professional.
\item Code review is important for knowledge sharing throughout an organization.
\item Automation is critical for scaling the process.
\item The code review itself provides a historical record.
\eit

Documentation:
\bit
\item Documentation is hugely important over time and scale.
\item Documentation changes should leverage the existing developer workflow.
\item Keep documents focused on one purpose.
\item Write for your audience, not yourself.
\eit

Testing Overview:
\bit
\item Automated testing is foundational to enabling software to change.
\item For tests to scale, they must be automated.
\item A balanced test suite is necessary for maintaining healthy test coverage.
\item "If you liked it, you should have put a test on it."
\item Changing the testing culture in organizations takes time.
\eit

Unit Testing:
\bit
\item Strive for unchanging tests.
\item Test via public APIs.
\item Test state, not interactions.
\item Make your tests complete and concise.
\item Test behaviors, not methods.
\item Structure tests to emphasize behaviors.
\item Name tests after the behavior being tested.
\item Don't put logic in tests.
\item Write clear failure messages.
\item Follow DAMP over DRY when sharing code for tests.
\eit

Test Doubles:
\bit
\item A real implementation should be preferred over a test double.
\item A fake is often the ideal solution if a real implementation can't be used in a test.
\item Overuse of stubbing leads to tests that are unclear and brittle.
\item Interaction testing should be avoided when possible: it leads to tests that are brittle because it exposes
implementation details of the system under test.
\eit

Larger Testing:
\bit
\item Larger tests cover things unit tests cannot.
\item Large tests are composed of a System Under Test, Data, Action, and Verification.
\item A good design includes a test strategy that identifies risks and larger tests that mitigate them.
\item Extra effort must be made with larger tests to keep them from creating friction in the developer workflow.
\eit

Deprecation:
\bit
\item Software systems have continuing maintenance costs that should be weighed against the costs of removing them.
\item Removing things is often more difficult than building them to begin with because existing users are often using
the system beyond its original design.
\item Evolving a system in place is usually cheaper than replacing it with a new one, when turndown costs are included.
\item It is difficult to honestly evaluate the costs involved in deciding whether to deprecate: aside from the direct
maintenance costs involved in keeping the old system around, there are ecosystem costs involved in having multiple
similar systems to choose between and that might need to interoperate. The old system might implicitly be a drag on
feature development for the new. These ecosystem costs are diffuse and difficult to measure. Deprecation and removal
costs are often similarly diffuse.
\eit

Version Control And Branch Management:
\bit
\item Use version control for any software development project larger than "toy project with only one developer that
will never be updated."
\item There's an inherent scaling problem when there are choices in "which version of this should I depend upon?"
\item One-Version Rules are surprisingly important for organizational efficiency. Removing choices in where to commit or
what to depend upon can result in significant simplification.
\item In some languages, you might be able to spend some effort to dodge this with technical approaches like shading,
separate compilation, linker hiding, and so on. The work to get those approaches working is entirely lost labor—your
software engineers aren't producing anything, they're just working around technical debts.
\item Previous research (DORA/State of DevOps/Accelerate) has shown that trunk-based development is a predictive factor
in high-performing development organizations. Long-lived dev branches are not a good default plan.
\item Use whatever version control system makes sense for you. If your organization wants to prioritize separate
repositories for separate projects, it's still probably wise for interrepository dependencies to be unpinned at
head/trunk based. There are an increasing number of VCS and build system facilities that allow you to have both small,
fine-grained repositories and a consistent "virtual" head/trunk notion for the whole organization.
\eit

Code Search:
\bit
\item Helping your developers understand code can be a big boost to engineering productivity. At Google, the key tool
for this is Code Search.
\item Code Search has additional value as a basis for other tools and as a central, standard place that all
documentation and developer tools link to.
\item The huge size of the Google codebase made a custom tool—as opposed to, for example, grep or an IDE's
indexing—necessary.
\item As an interactive tool, Code Search must be fast, allowing a "question and answer" workflow. It is expected to
have low latency in every respect: search, browsing, and indexing.
\item It will be widely used only if it is trusted, and will be trusted only if it indexes all code, gives all results,
and gives the desired results first. However, earlier, less powerful, versions were both useful and used, as long as
their limits were understood.
\eit

Build Systems And Build Philosophy:
\bit
\item A fully featured build system is necessary to keep developers productive as an organization scales.
\item Power and flexibility come at a cost. Restricting the build system appropriately makes it easier on developers.
\item Build systems organized around artifacts tend to scale better and be more reliable than build systems organized
around tasks.
\item When defining artifacts and dependencies, it's better to aim for fine-grained modules. Fine-grained modules are
better able to take advantage of parallelism and incremental builds.
\item External dependencies should be versioned explicitly under source control. Relying on "latest" versions is a
recipe for disaster and unreproducible builds.
\eit

Critique: Google's Code Review Tool:
\bit
\item Trust and communication are core to the code review process. A tool can enhance the experience, but it can't
replace them.
\item Tight integration with other tools is key to great code review experience.
\item Small workflow optimizations, like the addition of an explicit "attention set," can increase clarity and reduce
friction substantially.
\eit

Static Analysis:
\bit
\item Focus on developer happiness. We have invested considerable effort in building feedback channels between analysis
users and analysis writers in our tools, and aggressively tune analyses to reduce the number of false positives.
\item Make static analysis part of the core developer workflow. The main integration point for static analysis at Google
is through code review, where analysis tools provide fixes and involve reviewers. However, we also integrate analyses at
additional points (via compiler checks, gating code commits, in IDEs, and when browsing code).
\item Empower users to contribute. We can scale the work we do building and maintaining analysis tools and platforms by
leveraging the expertise of domain experts. Developers are continuously adding new analyses and checks that make their
lives easier and our codebase better.
\eit

Dependency Management:
\bit
\item Prefer source control problems to dependency management problems: if you can get more code from your organization
to have better transparency and coordination, those are important simplifications.
\item Adding a dependency isn't free for a software engineering project, and the complexity in establishing an "ongoing"
trust relationship is challenging. Importing dependencies into your organization needs to be done carefully, with an
understanding of the ongoing support costs.
\item A dependency is a contract: there is a give and take, and both providers and consumers have some rights and
responsibilities in that contract. Providers should be clear about what they are trying to promise over time.
\item SemVer is a lossy-compression shorthand estimate for "How risky does a human think this change is?" SemVer with a
SAT-solver in a package manager takes those estimates and escalates them to function as absolutes. This can result in
either overconstraint (dependency hell) or underconstraint (versions that should work together that don't).
\item By comparison, testing and CI provide actual evidence of whether a new set of versions work together.
\item Minimum-version update strategies in SemVer/package management are higher fidelity. This still relies on humans
being able to assess incremental version risk accurately, but distinctly improves the chance that the link between API
provider and consumer has been tested by an expert.
\item Unit testing, CI, and (cheap) compute resources have the potential to change our understanding and approach to
dependency management. That phase-change requires a fundamental change in how the industry considers the problem of
dependency management, and the responsibilities of providers and consumers both.
\item Providing a dependency isn't free: "throw it over the wall and forget" can cost you reputation and become a
challenge for compatibility. Supporting it with stability can limit your choices and pessimize internal usage.
Supporting without stability can cost goodwill or expose you to risk of important external groups depending on something
via Hyrum's Law and messing up your "no stability" plan.
\eit

Large-Scale Changes:
\bit
\item An LSC process makes it possible to rethink the immutability of certain technical decisions.
\item Traditional models of refactoring break at large scales.
\item Making LSCs means making a habit of making LSCs.
\eit

Continuous Integration:
\bit
\item A CI system decides what tests to use, and when.
\item CI systems become progressively more necessary as your codebase ages and grows in scale.
\item CI should optimize quicker, more reliable tests on presubmit and slower, less deterministic tests on post-submit.
\item Accessible, actionable feedback allows a CI system to become more efficient.
\eit

Continuous Delivery:
\bit
\item Velocity is a team sport: The optimal workflow for a large team that develops code collaboratively requires
modularity of architecture and near-continuous integration.
\item Evaluate changes in isolation: Flag guard any features to be able to isolate problems early.
\item Make reality your benchmark: Use a staged rollout to address device diversity and the breadth of the userbase.
Release qualification in a synthetic environment that isn't similar to the production environment can lead to late
surprises.
\item Ship only what gets used: Monitor the cost and value of any feature in the wild to know whether it's still
relevant and delivering sufficient user value.
\item Shift left: Enable faster, more data-driven decision-making earlier on all changes through CI and continuous
deployment.
\item Faster is safer: Ship early and often and in small batches to reduce the risk of each release and to minimize
time to market.
\eit

Compute As A Service:
\bit
\item Scale requires a common infrastructure for running workloads in production.
\item A compute solution can provide a standardized, stable abstraction and environment for software.
\item Software needs to be adapted to a distributed, managed compute environment.
\item The compute solution for an organization should be chosen thoughtfully to provide appropriate levels of
abstraction.
\eit

\chapter{Hands-On Machine Learning With Scikit-Learn And TensorFlow}

Frame The Problem:
\bit
\item Define the objective in business terms.
\item How will your solution be used?
\item What are the current solutions/workarounds (if any)?
\item How should you frame this problem (supervised/unsupervised, online/offline, etc.)?
\item How should performance be measured?
\item Is the performance measure aligned with the business objective?
\item What would be the minimum performance needed to reach the business objective?
\item What are comparable problems? Can you reuse experience or tools?
\item Is human expertise available? How would you solve the problem manually?
\item List the assumptions you (or others) have made so far and verify them if possible.
\eit

Get The Data:
\bit
\item List the data you need and how much you need.
\item Find and document where you can get that data.
\item Check how much space it will take.
\item Check legal obligations, and get access authorization if necessary.
\item Create a workspace (with enough storage space).
\item Get the data.
\item Make sure you data are not of insufficient quantity, poor quality, or non-representative of the population.
Especially validation and test set must be as representative as possible..
\item Convert the data to a format you can easily manipulate (without changing the data itself).
\item Ensure sensitive information is deleted or protected (e.g.\ anonymized).
\item Make a copy of the original dataset and work with this one.
\item Automate as much as possible, so you can easily get fresh data.
\eit

Explore The Data:
\bit
\item Create a Jupyter notebook to keep a record of your data exploration.
\item Study each attribute and its characteristics: name, type, percentage of missing values, noisiness and type of
noise, outliers, usefulness for the task basic statistics (mean, standard deviation, type of distribution, etc.).
\item For supervised learning tasks, identify the target attribute(s).
\item Sample a test set, put it aside, and never look at it (stratify if needed).
\item Visualize the data.
\item Study the correlations between attributes.
\item Study how you would solve the problem manually.
\item Identify extra data that would be useful.
\item Document what you have learned.
\item Try to get insights from a field expert for these steps.
\eit

Prepare The Data:
\bit
\item Write functions for all data transformations you apply, for the following reasons:
\item So you can easily prepare the data the next time you get a fresh dataset.
\item So you can apply these transformations in future projects.
\item To clean and prepare the test set.
\item To clean and prepare new data instances once your solution is live.
\item To make it easy to treat your preparation choices as hyperparameters.
\item Clean the data: fix or remove outliers, fill in missing values (e.g.\ with zero, mean, median, etc.) or drop their
rows (or columns).
\item Feature selection/engineering:
\item Drop the attributes that provide no useful information for the task.
\item Drop variables that have a very high percentage of missing values.
\item Drop variables that have a very low variation (i.e.\ not too much information).
\item Drop variables that have very low correlation with the target.
\item Find variables that are highly correlated with each other (i.e.\ same behaviour), and keep the ones that have
higher correlation with the target (drop the other ones).
\item Identify the promising transformations you may want to apply (e.g.\ combine existing features to produce more
useful ones or add promising transformations of features).
\item Select best features based on a metric. Either start with one variable and add more (forward selection), or start
with all variables and eliminate (backward elimination or recursive feature elimination).
\item Discretize continuous features.
\item Decompose features (e.g.\ categorical, date/time, etc.).
\item Standardize or normalize features (feature scaling). Machine learning algorithms don't perform well when the input
numerical attributes have very different scales. There are two common ways to get all attributes to have the same scale:
normalization and standardization. In normalization the values are shifted and rescaled so that they end up ranging from
0 to 1. We do this by subtracting the minimum value and dividing by the maximum minus the minimum. Standardization first
subtracts the mean value (so standardized values always have a zero mean), and then divides by the standard deviation so
that the resulting distribution has unit variance. Unlike normalization, standardization does not bound values to a
specific range, which may be a problem for some algorithms. However, standardization is much less affected by outliers.
\eit

Shortlist Promising Models:
\bit
\item If the data is huge, you may want to sample smaller training sets, so you can train many different models in a
reasonable time (be aware that this penalizes complex models such as large neural nets or Random Forests).
\item Train many quick-and-dirty models from different categories using standard parameters.
\item Measure and compare their performance.
\item For each model, use K-fold cross-validation and compute the mean and standard deviation of the performance measure
on the K folds.
\item Analyse the most significant variables for each algorithm.
\item Analyse the types of errors the models make. What data would a human have used to avoid these errors?
\item Shortlist the top three to five most promising models, preferring models that make different types of errors.
\eit

Fine-Tune The System:
\bit
\item You will want to use as much data as possible for this step, especially as you move toward the end of fine-tuning.
\item Fine-tune the hyperparameters using cross-validation.
\item Try ensemble methods. Combining your best models will often produce better performance than running them
individually.
\item Once you are confident about your final model, measure its performance on the test set to estimate the
generalization error.
\eit